---
title: 'Machine Learning Peer-graded Assignment: Prediction Assignment'
author: "Wen Xu"
date: "4/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ¨C a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from [the website here](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har).

## Goal
The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 


## Data

The training data for this project are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).

The test data are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv).


The data for this project come from this [source](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har). 

## Reference
Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 
Cited by 2 (Google Scholar)



## My Work

### Load the library, data and do some preprocessing 


```{r,results="hide"}
library(caret)
training = read.csv("pml-training.csv")
str(training)
```
From the result of *str(training)*, we found that the missing values are: **""**, **NA** and **"#DIV/0!"**. So let's reload the data.

```{r}
training = read.csv("pml-training.csv",na.strings = c("","NA","#DIV/0"))
testing = read.csv("pml-testing.csv",na.strings = c("","NA","#DIV/0"))
dim(training)
```
Also, we noticed that the first 7 columns are not related to the response *classe*. So we exclude them from both training and testing data sets.

```{r}
training = training[,8:ncol(training)]
testing = testing[,8:ncol(testing)]
```

### Cross Validation

There are enough observations in the training data set to separate them into a subtraining data and a subtesting data. 

```{r}
set.seed(11231)
intrain = createDataPartition(training$classe,p=0.5,list=F)
subtraining = training[intrain,]
subtesting = training[-intrain,]
```

### Dealing with missing value columns
```{r}
NAs = apply(subtraining,2,function(x){sum(is.na(x))})
table(NAs)
```
For all the columns, either it does not contain a missing value, or it contains **9614** missing values! Note that each column contains only **9812** observations, which is to say, columns with missing values miss the most of data -- so we throw those columns away rather than imputing.

```{r}
subtesting = subtesting[,colSums(is.na(subtraining))==0]
testing = testing[,colSums(is.na(subtraining))==0]
subtraining = subtraining[,colSums(is.na(subtraining))==0]
```
Keep in mind that all data processing should be done in the same way to both training and testing.

### Model1: **boosting** model using **gbm** 



```{r,results="hide"}
library(gbm)
```
```{r}
model = gbm(classe~.,data=subtraining,distribution='multinomial',n.trees=4000)
pred = predict(model,newdata=subtraining,n.trees=4000,type='response')
predict = as.factor(colnames(pred)[apply(pred,1,which.max)])
confusionMatrix(predict,subtraining$classe)$overall[1]
```
Only more than 70% accuracy -- and this is the training accuracy! Let's check the subtesting set accuracy.

```{r}
pred = predict(model,newdata=subtesting,n.trees=4000,type='response')
predict = as.factor(colnames(pred)[apply(pred,1,which.max)])
confusionMatrix(predict,subtesting$classe)$overall[1]
```


### Model2: **random forest** model using **randomForest** 

```{r,results="hide"}
library(randomForest)
```
```{r}
model_rf = randomForest(classe~.,data=subtraining,type="class")
confusionMatrix(predict(model_rf,subtraining),subtraining$classe)$overall[1]
```
The training accuracy is full. Let's check the subtesting set accuracy.

```{r}
confusionMatrix(predict(model_rf,subtesting),subtesting$classe)$overall[1]
```
Still high -- so the out of sample error is very small -- so we choose this random forest model to predict on the testing set.
```{r}
prediction = predict(model_rf,testing)
```

## Conclusion

The random forest model seems perfect for this data and our prediction on the test data are: 
```{r,echo=FALSE}
prediction
```











